On the Functionality of the Multilayer Perceptron

### 2020-09-24 Thesis/Sections/Methodology.tex
Updatess to Methodology section

% ================================
% Landon Buell
% Kevin Short
% Physics 799%
% 30 July 2020 
% ================================

@@

\section{The Neural Network}
\label{sec-TheNeuralNetwork}

@@

\subsection{An Introduction to Neural Networks}
\label{sec-NerualNetworkIntro}

\paragraph*{}The task of algorithmically matching sound files to musical instrument classes in exceedingly difficult through conventional computer programming techniques. Because of the complexity of sound waves, brain's interpretation them, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an architecture that allows for a computer to \textit{learn} and change and update itself as it is presented with more data. Such an algorithm exists in the form of a \textit{Neural Network} \cite{Geron2,Goodfellow,Levine}.

\paragraph*{}A neural network is a mathematical function that seeks to produce an approximation $F^*$, of some usually unknown function $F$ . For audio recognition, $F$ is some function, or series of functions, that occurs in the brain which allows for a listener to take the sensation of audio, and map it to a known musical instrument. For a neural network, a model must construct an approximate function $F^*$ using a set of parameters $\Theta$, which allows the model to map a series of inputs, $x$ (called \textit{features}) to a series of outputs, $y$ (called \textit{predictions}) \cite{Goodfellow,James,Virtanen}.

\paragraph*{}Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result of such an analogy is a computer program that is reminiscent of the brain.

\paragraph*{}Over the course of their lives, humans will learn to map sounds to sources almost effortlessly. Given an example, and the appropriate label, humans can do this with very reasonable accuracy over a wide array of sounds \cite{Olsen,White}. We can simplify the idea of humans recognizing sound as some function or operation that occurs within the brain, accepting an input sound, and produces an output label. Similarly, a neural network can be constructed, presented with \textit{features} of multiple, labeled sound waves and learns a set of parameters $\Theta$ that allows for the mapping to a source.

\paragraph*{}For this particular project, a neural network has been constructed to perform a \textit{classification} task. A classification task involves mapping the input data, $x$ to one of $k$ possible \textit{classes}, each represented by an integer, $0$ through $k-1$. Each of the $k$ classes represents a particular musical instrument that could have produced the sound-wave in the audio file.

@@

\subsection{The Structure of a Neural Network}
\label{sec-NetworkStructre}

\paragraph*{}A Neural Network is simply a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes some real input, typically an array of real double-precision floating-point numbers, and returns a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. It is this process of transforming inputs successively in a particular order until an output is attained \cite{Geron,Loy}. This output encodes the models final "decision" given a unique input.

\paragraph*{}Other than inspiration from the brain, Neural Networks are name \textit{networks} because of their nested composition nature. The model can be represented by a linear or acyclic computational graph which successively maps exactly how the repeated composition is structured. In a \textit{feed-forward network}, information is passes successively in one direction. For example functions could be chained together such as \cite{Goodfellow}
\begin{equation}
\label{eqn-FunctionChain}
F(x) = f^{(L-1)}\big[f^{(L-2)}\big[...f^{(1)}\big[f^{(0)}\big[x]\big]\big]...\big]\big]
\end{equation}
Where each function $f^{(l)}$ represents a layer of the network model. The number of layers in a neural network is referred to as the network \textit{depth}. The dimensionality of each layer is referred to as the network \textit{width} \cite{Geron,Loy}.

\paragraph*{}A network model that contains $L$ unique layers is said to be an $L$-Layer Neural Network, with each layer usually indexed by a superscipt, $0$ through $L-1$. Layer $0$ is said to be the \textit{input layer} and layer $L-1$ is said to be the \textit{output layer}. The function that represents a layer $(l)$ is given by 
\begin{equation}
\label{eqn-LayerFunction}
f^{(l)} : x \in \mathbb{R} \rightarrow y \in \mathbb{R}
\end{equation}
The value of $x$ can also be index by layer, we call the array $x^{(l)}$ the \textit{activations} of layer $l$ \cite{Goodfellow,Loy}.

\paragraph*{}The model is recursive by nature, the activations from one layer, $l-1$, are used to directly produce the activations of the next successive layer $l$. Thus eqn. (\ref{eqn-LayerFunction}) can be alternatively written as:
\begin{equation}
\label{eqn-altLayerFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R} \rightarrow x^{(l)} \in \mathbb{R}
\end{equation}
The array of activations, $x^{(0)}$, is the raw input given the neural network, most commonly called \textit{features}. Conversely, the activations $x^{(L-1)}$ are commonly called the network \textit{output} \cite{Geron,James,Loy}.

\begin{algorithm}[H]
\caption{Forward propagation system in a standard deep neural network. Each iteration in the main \textit{for-loop} represents the execution of a layer, and passing the result to the "next" layer function. A practical application of this algorithm should include batches of samples instead of a single sample.
\textcolor{red}{I will put this some place better soon! It is here for future placement}}
\label{algFeedForward}

\begin{algorithmic}
\REQUIRE Network with $L$ layers - $0,1,2,....,L-2,L-1$
\REQUIRE Set of weighting parameters - $W^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of bias parameters - $b^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of layer activation function - $a^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Input sample - $x^{(0)}$

$Z = \{\} \leftarrow$ array to hold pre-activations \\
$X = \{\} \leftarrow$ array to hold final activations

\FOR {$l = 1,2,...,L-2,L-1$}
	\item Compute pre-nonlinear activation
	\item $z^{(l)} \leftarrow W^{(l)}x^{(l-1)} + b^{(l)}$
	\item Apply activation function
	\item $x^{(l)} \leftarrow a^{(l)}[z^{(l)}]$
	\item Store values for later use in optimizer
	\item $Z \leftarrow Z.add(z^{(l)})$
	\item $X \leftarrow X.add(x^{(l)})$
\ENDFOR

\item Network final prediction is last object in 'act' array.
\item $y^* \leftarrow x^{(L-1)} = X[-1]$
\item Return the pre-nonlinear activations and final-nonlinear activations
\RETURN $Z , X$

\end{algorithmic}
\end{algorithm}

@@

\subsection{Layers Used in Classification Neural Network}
\label{sec-Layers}

\paragraph*{}As stated previously, a neural network is composed of a series of functions that are called successively to transform features (an input) into a prediction (an output). As shown in eqn. (\ref{eqn-altLayerFunction}), each function feeds directly into the next as to form a sort of computational graph \cite{Goodfellow}.

\paragraph*{}Typically, a layer function can be divided into two portions: (i.) a Linear transformation, with a bias addition, and (ii.) an element-wise activation transformation. This activation function typically is a non-linear function which allows for the modeling of increasingly complex decision-boundaries. Step (i.) is usually in the form of a matrix-vector equation:
\begin{equation}
\label{eqn-matVectEqn}
z^{(l)} = W^{(l)} x^{(l-1)} + b^{(l)}
\end{equation}
Where $W^{(l)}$ is the \textit{weighting-matrix} for layer $l$, $b^{(l)}$ is the \textit{bias-vector} for layer $l$, $z^{(l)}$ are the \textit{linear-activations} for layer $l$ and $x^{(l-1)}$ is the final activations for layer $l-1$.
Step (ii.) is usually given by some \textit{activation function}:
\begin{equation}
\label{eqn-elementActivation}
x^{(l)} = \sigma^{(l)} \big[ z^{(l)} \big]
\end{equation}
Where $x^{(l)}$ is the final activations for layer $l$ and $z^{(l)}$ is given in equation (\ref{eqn-matVectEqn}). $\sigma^{(l)}$ is some activation function, which is often use to enable the modeling of more complex-decision boundaries.

\paragraph*{}The combination of eqn. (\ref{eqn-matVectEqn}) and eqn. (\ref{eqn-elementActivation}), for a layer $l$ , make up the function represented by that layer, as in eqn. (\ref{eqn-altLayerFunction}). Below, we discuss and describe the types of layer functions that are used to produce the classification model in this project.

% ================

\subsubsection{Dense Layer}
\label{subsec-DenseLayer}

\paragraph*{}The Linear Dense Layer, often just called a \textit{Dense} Layer, or \textit{Fully-Connected} layer for short, was one of the earliest function types used in artificial neural network models \cite{Geron2,Loy}. They are among the most commonly used layer types. A dense layer is composed of a layer of \textit{artifical neurons}, each of which holds a numerical value within it, called the \textit{activation} of that neuron. This idea was developed back from McCulloch and Pitts' work \cite{McCulloch}, and was expanded upon by Frank Rosenblatt in  1957 \cite{Geron}.

\paragraph*{}We model a layer of neurons as a vector of floating-point numbers. Typically, it is required that the array be one-dimensional. Suppose a layer $(l)$ contains $n$ artificial neurons. We denote the array that hold those activations as $x^{(l)}$ and is given by:
\begin{equation}
\label{layer-DenseNeurons}
\vec{x}^{(l)} = \Big[ x_0, x_1, x_2, \hdots , x_{n-2}, x_{n-1} \Big]^T
\end{equation}
The activation of each entry is given by a linear-combination of activations from the previous layer, as outlined in eqn.(\ref{eqn-matVectEqn}) and eqn.(\ref{eqn-elementActivation}).

\paragraph*{}Suppose the layer $l-1$ contains $m$ neurons. Then the weighting-matrix, $W^{(l)}$ has shape $m \times n$, the bias-vector $b^{(l)}$ has shape $m \times 1$. We can denote the function in a similar manner to eqn.(\ref{eqn-altLayerFunction}) as:
\begin{equation}
\label{eqn-FunctionDense}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(m \times 1)} \rightarrow x^{(l)} \in \mathbb{R}^{(n \times 1)}
\end{equation}

Thus for a dense layer $l$, the exact values of each activation is given by \cite{Geron,Loy}:
\begin{equation}
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{n-1}
\end{bmatrix}^{(l)} =
\sigma^{(l)} \Bigg(
\begin{bmatrix}
w_{0,0} & w_{0,1} & \hdots & w_{0,m-1} \\
w_{1,0} & w_{1,1} & \hdots & w_{1,m-1} \\
\vdots & \vdots & \ddots & \vdots  \\
w_{n-1,0} & w_{n-1,1} & \hdots & w_{n-1,m-1} 
\end{bmatrix}^{(l)} 
\begin{bmatrix}
x_0 \\ x_1 \\ \vdots \\ x_{m-1}
\end{bmatrix}^{(l-1)} +
\begin{bmatrix}
b_0 \\ b_1 \\ \vdots \\ b_{n-1}
\end{bmatrix}^{(l)} \Bigg)
\end{equation}
or more compactly:
\begin{equation}
\label{eqn-DenseFeedForward}
x^{(l)} = \sigma^{(l)} \Big( W^{(l)} x^{(l-1)} + b^{(l)} \Big)
\end{equation}
Eqn. (\ref{eqn-DenseFeedForward}) is generally referred to as the \textit{dense layer feed-forward equation} \cite{Goodfellow}.

% ================

\subsubsection{2-Dimensional Convolution Layer}

\paragraph*{}Convolution layers emerged from studying the brain's visual cortex, and have been used from image recognition related tasks for around 40 years. \cite{Geron,Loy}). This layer get's it's name from it's implementation of the mathematical \textit{convolution} function. The 2D discrete-convolution of function or 2D array $A[i,j]$ and $B[i,j]$ is given by \cite{Goodfellow}:
\begin{equation}
\label{eqn-convolution}
C[i,j] = (A * B)[i,j] = \sum_{u}\sum_{v} A[i,j]B[i - u,j - v]
\end{equation}
Note that convolution is commutative: $(A * B) = (B * A)$. We implement a convolutional layer $f^{(l)})$, by creating a series of $K$ weighting matrices, called \textit{filters}, each of size $m \times n$, where $m,n \in \mathbb{Z}$. Often we choose $m = n$ and call the shape the \textit{convolution kernel size} \cite{Loy, Goodfellow}.

\paragraph*{}In the case of a 2D input array, $x^{(l-1)} \in \mathbb{R}^{(N \times M)}$, the convolutions filters "step" through the input data and repeatedly compute the element-wise product of each $m \times n$ weighting kernel, and the local activation of the $x^{(l-1)}$ array. The step size in each of the 2-dimension is known as the \textit{stride}. Each of the $K$ filters is used to generate a $m \times n$ feature map from the input activation. For an appropriately trained network, some filters may be dedicated to detecting vertical lines, horizontal lines, sharp edges, areas of mostly one color, etc. \cite{Loy,Geron}.

\paragraph*{}In general, for an $N \times M$ input, with kernel size $m \times n$, with stride size $1 \times 1$, and $K$ feature maps.
\begin{equation}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(N \times M)} \rightarrow
x^{(l)} \in \mathbb{R}^{([N-n+1] \times [M-m+1] \times K)}
\end{equation}
As an example, consider a single filter map over input $x \in \mathbb{R}^{(4 \times 3)}$ and filter map $W \in \mathbb{R}^{(2 \times 2)}$ as shown in fig. (\ref{fig-2DConvExample}):
\begin{figure}[H]
\label{fig-2DConvExample}
\begin{center}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
\caption{Input Activations}
\end{subfigure}
\begin{subfigure}{0.45\textwidth}
\includegraphics[scale=0.6]{../figures/2DConvExample(b)}
\caption{Convolution Filter}
\end{subfigure}
\begin{subfigure}{0.9\textwidth}
\centering
\includegraphics[scale=0.6]{../figures/2DConvExample(c)}
\caption{Convolution Result}
\end{subfigure}
\end{center}
\caption{The result of convolving an input (a) with an filter map (b) is a new set of activations (c). This Image was adapted from Goodfellow, pg. 325 \cite{Goodfellow}}
\end{figure}
Note that formal implementations include a bias vector and an activation function.

\paragraph*{}For 2D convolution over input $x^{(l-1)}$ with feature map $W^{(l)}_k$, producing activations $x^{(l)}$, we compute the activations $x^{(l)}_{k}$ as \cite{Goodfellow}
\begin{equation}
\label{eqn-ConvFeedForward}
x^{(l)}_k = \sigma^{(l)}\bigg[ \Big( \sum_{u=0}^{m-1} \sum_{v=0}^{n-1} x^{(l-1)}(i,j) W^{(l)}_k(i - u,j - v) \Big) + b^{(l)}_k \bigg]
\end{equation}
\textcolor{red}{Please check my math here!}
This operation repeats for each of the $K$ feature maps. Each maps has it's own $n \times m$ weighting matrix and appropriately shaped bias .

\paragraph*{}The convolution layers allows for several advantages. Among these are (i) \textit{sparse-connectivity}: not every single activation (pixel) is connect to every single output pixel, so it is more computationally efficient, (ii) \textit{positional invariance}: key features can be identified regardless of where they are in the layer

% ================

\subsubsection{2-Dimensional Maximum Pooling Layer}

\paragraph*{}A Maximum Pooling layers simply returns the maximum neuron activation in a group of neurons. In the case of 2D Max Pooling, we choose a kernel size to be $m \times n$, just like in the 2D Convolution layer, and extract the maximum value in each window, and then step along accrdoing to the chosen stride size \cite{Loy,Goodfellow}.As an example, consider again figure (\ref{fig-2DConvExample}). Using the $2 \times 2$ kernel on the $3 \times 4$ input, each box would then contain the maximum value of the input activations. We detail this in fig. (\ref{fig-2DMaxPool}):
\begin{figure}[h]
\label{fig-2DMaxPool}
\begin{center}
\includegraphics[scale=0.6]{../figures/2DConvExample(d)}
\end{center}
\caption{The result of 2D maximum-pooling an input array. This image was adapted from Loy, pg. 126 \cite{Loy}}
\end{figure}

\paragraph*{}Pooling layers, such as maximum pooling, average pooling, or similar are typically placed after a layer or a set of layers of convolution. The purpose of this arrangement is to reduce the number of weights, thereby dropping the complexity of the model and ensuring only features with large activation values are preserved \cite{Geron,Loy,Goodfellow}.

% ================

\subsubsection{1-Dimensional Flattening Layer}

\paragraph*{}A flattening layer is used to compress an array with two or more dimensions down into a single dimension. For a flattening layer $l$, multidimensional activations in layer $l-1$ are compressed down into a single dimensional array. We can use function notation to express this as:
\begin{equation}
\label{eqn-FlattenFunction}
f^{(l)} : x^{(l-1)} \in \mathbb{R}^{(M \times N \times ....)} \rightarrow
x^{(l)} \in \mathbb{R}^{(MN...\times 1)}
\end{equation}
The numerical value of each activation is left unchanged. For a layer with activation shape$N \times M$, the resulting activations are reshaped into $NM \times 1$ as shown in eqn (\ref{eqn-FlattenFunction}).

\paragraph*{}Flattening Layer are most commonly used to prepare activations for entry into dense layer or series of dense layers. For example, 2D or 3D images are typically processed with 2D convolution, which may output a 2D or 3D array of activations. Those values are then flattened to 1 dimensions, which can then be passed into dense layers for further processing.

% ================

\subsubsection{1-Dimensional Concatenation Layer}

\paragraph*{}The 1-Dimensional Concatenation Layer, also called 1D-Concat layer takes separate vectors of activations

@@

\subsection{Activation Functions Used in Network Layers}
\label{sec-ActivationFunctions}

\paragraph*{}Activation functions are a key parameter in the behavior of neural networks. As discussed in sec. (\ref{sec-Training}) a layer function, $f^{(l)}$ is generally composed of a linear transformation, as in eqn. (\ref{eqn-matVectEqn}), and then an element-wise activation function as in eqn. (\ref{eqn-elementActivation}). It is this second step that allows for neural networks to model the incredibly complex decision boundaries that are found in real-world problems \cite{Geron,Loy}. In this section, we detail the activation functions used in out classification neural network.

\subsubsection{Rectified Linear Unit}

\paragraph*{}The Rectified Linear Unit (ReLU) activation function acts element-wise on the activations in a given layer. If the activation of a neurons is non-zero or non-negative, the value is untouched, otherwise a $0$ is returned. For an input activation $x$, ReLU is defined by:
\begin{equation}
\label{eqn-ReLU}
\text{ReLU}\big[x\big] = \max\big(0,x\big) = \left\{
\begin{array}{ll}
    x & \text{if\hspace{0.5cm}} x > 0 \\
    0 & \text{otherwise}
\end{array} 
\right.
\end{equation}
ReLU is only defined on a real domain. We provide a visualization of the function in fig. (\ref{fig-ReLU}).
\begin{figure}[H]
\label{fig-ReLU}
\begin{center}
\includegraphics[scale=0.2]{../Figures/RectifiedLinearUnit}
\end{center}
\caption{Rectifed Linear Unit (ReLU) activation function}
\end{figure}

\subsubsection{Softmax}

@@

\subsection{Training a Neural Network}
\label{sec-Training}

\paragraph*{}A neural network's purpose is to produce a function $F^*$ that approximates an unknown function $F$, using a set of parameters, $\Theta$. The model must have a procedure for updating the parameters in $\Theta$ to allow for a reasonably close approximation of $F$. To better understand this, we turn to Tom Mitchell's explanation of a learning algorithm \cite{Goodfellow,Mitchell}:
\begin{quote}
A computer program is said to learn from experience $E$ with respect to some class of tasks $T$ and performance measure $P$, if it's performance at tasks in $T$, as measured by $P$, improves with experience $E$.
\end{quote}
Without any human intervention, a model must update itself to improve it's performance at a give task a new information is presented to it. To do this, the model must be constructed with a training procedure in mind.

\paragraph*{}We consider the set of parameters $\Theta$ as a concatenation of each applicable layer's weighting matrix and bias vector such that:
\begin{equation}
\label{eqn-Theta}
\Theta = \big\{
W^{(0)}, b^{(0)}, W^{(1)}, b^{(1)}, W^{(2)}, b^{(2)}, ... ,W^{(L-2)}, b^{(L-2)}, W^{(L)}, b^{(L-1)},
\big\}
\end{equation}
Each parameter is floating-point number, which is stored in a weight matrix $W^{(l)}$ or bias vector $b^{(l)}$. Each value for each parameter contributes to the output of the function. Note that $W^{(l)}$ or $b^{(l)}$, itself is a parameter to train, but rather the entries that make up those structures must be trained. In increasingly complicated neural networks, there can be upwards of hundreds of thousands,or even millions of elements in $\theta$ making a neural network a functions that exist is a very high dimensional parameter-space \cite{Geron,Goodfellow}.

\subsubsection{The Cost Function}

\paragraph*{}Suppose we pass a training sample, given by the feature-vector $x^{(0)}$, with an expected outcome given by the vector $y$, into the neural network. After the network finishes processing, it's output activations are given by the vector $x^{(L-1)}$, also noted at $y^*$. For a reasonably trained model, we expect the vector $y^*$ to be "similar" to $y$, indicating that with the provided data, the network has made a valid prediction. Conversely, for an untrained network, $y^*$ is not likely to be "similar" to $y$ at all, indicating that the network has made a poor prediction.

\paragraph*{}To quantify the difference between the model's prediction, $y^*$ and the expected output, $y$, we introduce a \textit{cost function}, $J(y^*,y)$, to the network \cite{Goodfellow,James}. The cost function, also called a \textit{loss} or \textit{objective} function, compares the neural networks output, $y^*$ and the expected output $y$. It returns a single value, usually a floating-point number which measures the \textit{poorness} of the prediction. A high cost value indicates a \textit{poor} prediction, and a low cost indicates a reasonable prediction.

\paragraph*{}$J(y^*,y)$ can take many forms and is often dependent on particular task or data set provided \cite{James}. For this $k$-classes classification task, we choose to use a \textit{Categorical Crossentropy} (CXE) cost function. For a single predictions sample, it is defined \cite{Goodfellow}:
\begin{equation}
\label{eqn-CategoricalCrossentropy}
\text{CXE}\big[ y^* , y \big] = - \sum_{i=0}^{k-1} y^*_i \log_{10}(y_i) 
\end{equation}
In practice, this operation is performed over a batch of samples and an average is computed and used a \textit{batch loss} \cite{Goodfellow}.

\subsubsection{Gradient Based Learning}

\paragraph*{}We have developed a method for quantifying the difference between a given output $y^*$ and an expected output $y$ with the inclusion of a cost function. For a trained neural network, we expect that samples consistently produce a low cost value, particularly on previous unseen inputs, which indicates that the model has appropriately \textit{generalized}. The process of training the model is to choose the parameters in $\Theta$ that allows for this behavior of the cost function. We then treat the training process of a neural network as a higher dimensional \textit{optimization} problem.

\paragraph*{}A neural network uses \textit{indirect optimization}, which contrasts from pure optimization. Deep Learning expert Ian Goodfellow described this \cite{Goodfellow}:
\begin{quote}
In most learning scenarios, we care about some performance measure $P$,... We reduce a different cost function, $J(\theta)$ in the hope that doing so will also imprvce $P$.
\end{quote}
By choosing an appropriate cost function, and selecting the parameters $\Theta$ that minimize the average cost over a data set, we also assume that doing so minimized the error in the performance $P$.

\paragraph*{}The cost of a sample is dependent on training labels $y$ and the network output $y^*$. The output is given by the final layer activation $x^{(L-1)}$, which in turn are produced by the previous layer and so forth, as demonstrated in eqn. (\ref{eqn-altLayerFunction}), This recursive nature combined with the dimensionality of the parameter object $\Theta$ makes an analytical solution to the optimization impractical \cite{Geron,Goodfellow,James}. We instead optimize with a gradient-based method.

\paragraph*{}We can reduce the cost given a set of paramaters, $J(\Theta)$, by moving each element in $\Theta$ with small steps in the direction of the negative partial derivative of each parameter. In the case of the high-dimensional $\Theta$ object, we compute the partial derivative with respect to each weight and bias in the form of the \textit{gradient vector}. We denote the gradient of $J$ with respect to the parameters in $\Theta$ as:
\begin{equation}
\label{eqn-CostGradient}
\nabla_{\Theta}\big[ J \big] = 
\Big[ 	\frac{\partial J}{\partial W^{(0)}}, \frac{\partial J}{\partial b^{(0)}}, 
		\frac{\partial J}{\partial W^{(1)}}, \frac{\partial J}{\partial b^{(1)}}, \hdots ,
		\frac{\partial J}{\partial W^{(L-1)}}, \frac{\partial J}{\partial b^{(L-1)}} \Big]^T
\end{equation}
Where $\frac{\partial J}{\partial W^{(l)}}$ is taking the partial derivative of each element in the $W^{(l)}$ matrix, and preserves the shape.

\paragraph*{}Due to the nested composition of the network output in eqn. (\ref{eqn-FunctionChain}), and subsequently the cost itself, we must use the chain rule of calculus to work backwards through the neural network to compute the partial derivative with respect to each parameter in $\Theta$. The process of working backwards to compute each element in the gradient vector is called \textit{back-propagation} \cite{Geron}. Below we detail an algorithm in pseudo-code to back-propagate through a fully-connected nueral network, such as the multilayer perceptron outlines in sec.(\ref{sec-NetworkStructre})

\begin{algorithm}[H]
\caption{Backwards propagation system, in a standard densely connected deep neural network. Each iteration in the \textit{for-loop} computes the gradient of the cost function $J$ with respect to the weight and bias arrays. Each element in those arrays is then the discrete gradient of that parameter. A practical application of this algorithm should include batches of samples instead of a single sample}
\label{algBackProp}

\begin{algorithmic}
\REQUIRE Cost/Objective function $J$ and learning rate $\alpha$
\REQUIRE Set of weighting parameters - $W^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of bias parameters - $b^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of layer activation function - $a^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of layer activation function derivatives - $\partial a^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of pre-nonlinear activation - $Z^{(i)}, i \in \{0,1,....,L-1\}$
\REQUIRE Set of post-nonlinear activation - $X^{(i)}, i \in \{0,1,....,L-1\}$

Execute forward pass in algorithm (\ref{algFeedForward}) and compute the gradient of the cost with respect to the final layer activations \\
$dx \leftarrow \nabla_{(y^*)}J(y,y^*)$ \\
Intialize $\nabla J$ as output object, should have same shape as $\Theta$

\FOR {$L-1,L-2,...,2,1$}
	\item Compute gradient w.r.t pre-nonlinear activation portion of layer function
	\item $dx^{(l)} \leftarrow \nabla_{Z^{(l)}}J = dx^{(l)} \odot \partial a^{(l)}[ Z^{(l)} ]$
	\item Compute gradient w.r.t weighting and bias elements
	\item $db \leftarrow \nabla_{b^{(l)}}J = dx^{(l)}$
	\item $dW \leftarrow \nabla_{W^{(l)}}J = dx^{(l)} \cdot X^{(l-1)}$

	\item Add $db$ and $dW$ steps to $\nabla J$ object
	\item $\nabla J = \nabla J .Add(dW,db)$
\ENDFOR

\item Return gradient w.r.t to each parameter in $\Theta$
\RETURN $\nabla J$

\end{algorithmic}
\end{algorithm}

\paragraph*{}After (i) computing the gradient, we can scale is by a desired learning rate, $\alpha$ and (ii) add it element-wise to the exisitint elements in $\Theta$. By repeating steps (i) and (ii) in sucession, we gradually drive the cost function to produces consistently lower and lower values. This is called \textit{gradient descent} and is the basis for many optimization algorithms. We show the general update rule on iteration $(i)$ for gradient based learning over a batch of $m$ samples in eqn. (\ref{eqn-GradientLearning}).
\begin{equation}
\label{eqn-GradientLearning}
\Theta^{(i)} = \Theta^{(i-1)} + (-\alpha)\frac{1}{m} \nabla_{\Theta} \Big[\sum_{n=1}^{m} J(y^{*(n)},y^{(n)}) \Big] \\
\end{equation}

\subsubsection{The Optimizer}

\paragraph*{}For this project, we employ an \textit{Adaptive-Moments} optimization algorithm, also called \textit{ADAM} for short. It is a powerful variant of algorithms with an adaptive learning rate. It tracks an exponentially decaying average of past gradients, as well as exponentially decaying average of past squared gradients \cite{Geron}. This produces a far more aggressive optimizer at a higher computational cost. For a given iteration $(i)$, The ADAM update is given:
\begin{equation}
\label{eqn-ADAMupdate}
\begin{split}
s^{(i)} &= \rho_1 s^{(i-1)} + (1 - \rho_1) \frac{1}{m} \nabla_{\Theta} \Big[\sum_{n=1}^{m} J(y^{*(n)},y^{(n)}) \Big]  \\
r^{(i)} &= \rho_2 r^{(i-1)} + (1 - \rho_2) \frac{1}{m} \nabla_{\Theta} \Big[\sum_{n=1}^{m} J(y^{*(n)},y^{(n)}) \Big] \odot
							\frac{1}{m} \nabla_{\Theta} \Big[\sum_{n=1}^{m} J(y^{*(n)},y^{(n)}) \Big]  \\
s'^{(i)} &= \frac{s^{(i)}}{1-\rho_1^t} \\
r'^{(i)} &= \frac{r^{(i)}}{1-\rho_2^t} \\
\Theta^{(i)} &= \Theta^{(i-1)} + (-\alpha)\frac{s'^{(i)}}{\sqrt{r'^{(i)}}+\delta} \\
\end{split}
\end{equation}

\paragraph*{}ADAM has experimentally shown to be a very powerful and robust optimizer useful for a wide range of tasks. Because of the two decay constants $\rho_1$ and $\rho_2$, we an compound and accumulate the values of past gradients to continue to push the cost to lower and lower values, even if the magnitude of the gradient becomes very small \cite{Geron}.

\begin{algorithm}[H]
\caption{Adaptive-Moments (ADAM) optimizer for a neural network}
\label{algAdaGrad}

\begin{algorithmic}
\REQUIRE Step size $\alpha$
\REQUIRE Small constant $\delta$ for numerical stabilization, usually about $10^{-7}$.
\REQUIRE constants $\rho_1$, $\rho_2$ used got exponential decay rates, usually $0.9$ and $0.999$ respectively.
\REQUIRE Subroutine/function to compute gradient of cost function.
\REQUIRE Mini-batch size, $m$
\REQUIRE Stopping criterion $S$

Initialize moment variables and iteration counter $s = 0$, $r = 0$, $i = 0$ \\

\WHILE {Stopping Criterion $S$ is \FALSE}
	\item Extract a mini-batch of $m$ samples from larger data set $X$. $[x^{(0)},x^{(1)},...,x^{(m-1)}]$ and corresponding target values 
	$[y^{(0)},y^{(1)},...,y^{(m-1)}]$.
	\item Compute numerical gradient estimate of each sample in batch. This can be done with standard back-propagation in algorithm (\ref{algBackProp}) and 			 	normalize by batch size $m$:
	\item $\nabla J \leftarrow \frac{1}{m} \nabla_{\Theta} \Big[\sum_{n=1}^{m} J(y^{*(n)},y^{(n)}) \Big]$

	\item Compute first bias moment: $s \leftarrow \rho_1 s + (1 - \rho_1) \nabla J$
	\item Compute second bias moment: $r \leftarrow \rho_2 s + (1 - \rho_2) \nabla J \odot \nabla J$
	\item First bias correction: $s' \leftarrow \frac{s}{1 - \rho_1^i}$
	\item Second bias correction: $r' \leftarrow \frac{r}{1 - \rho_2^i}$

	\item Compute And Apply update:	$\Delta \Theta = (-\alpha) \frac{s'}{\sqrt{r'} + \delta}$
	\item $\Theta = \Theta + \Delta \Theta$	
	\item Update Iteration number: $i \leftarrow i + 1$
\ENDWHILE

\end{algorithmic}
\end{algorithm}

@@

\subsection{Chosen Model Architecture}
\label{sec-Architecture}

>>>>

% ================================
% Landon Buell
% Kevin Short
% Physics 799%
% 24 Sept 2020 
% ================================

@@

\section{Methodology}
\label{sec-Methodology}

\paragraph*{}We outline and detail the steps taken to execute this project from beginning to end.

@@

\subsection{Designing the Function}

\paragraph*{}Consider the biological process of hearing a sound wave and matching it to a source. We can model this behavior by some unknown function $F$. We produce an approximation of that function $F^*$ that can map the contents of a sound file that have been generated by a chaotic synthesizer to a potential source. For a set of inputs $\vec{x} = \big\{ x_0 , x_1 , x_2 , ... , x_{p-1} \big\}$ and a set of classes $0$ through $k-1$, we denote this function as:
\begin{equation}
\label{eqn-MappingFunction}
F^*: \vec{x} \rightarrow \big\{ 0 , 1 , 2 , ... , k-1 \big\}
\end{equation}

@@

\subsection{Collecting and Pre-processing Raw Data}

\paragraph*{}In order to train the Multimodal neural network to identify musical instrument sources, we need a suitable data set to present to the model. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \textcolor{blue}{Citations!}. These contain short segments of musical instruments performing a single note or a collection of notes in succession.

\paragraph*{}To ensure that these data sets are roughly homogeneous, we read each sample from it's original format, \textit{.aif}, \textit{.mp3}, or similar, and rewrite each sample as a new \textit{.wav} files, sampled at 44.1 kHz, with a 16 bit depth \textcolor{red}{check this!}. This ensures that all data will have a consistent format when features are extracted. We also use this stage to ensure that each audio file has a correct label, and to determine the number of unique output classes.

@@

\subsection{Designing Classification Features}

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of classification features. These are properties of wave forms that can be represented by a numerical value or several numerical values and are used as the primary tool in classification. There are used in place of a full waveform to represent a file's contents We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\subsection{Designing A Complementary Network Architecture}

\paragraph*{}With the appropriate set of features designed, and the number of output determined, we can organize the structure of the neural network function. We construct a multimodal neural network than processes two input arrays derived from the same audio sample that share a common label. This network is designed to handle and process each respective input independently, and concatenate the results to produce a single output.

@@

\subsection{Testing and Evaluating Network Performance}

\paragraph*{}We divide the raw data set up into a training and testing sets. We employ cross-validation and compute the results of performance metrics to ensure that out model is making reliable predictions, and generalize appropriately. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance.

@@

\subsection{Running Predictions of Chaotic Synthesizer Files}

\paragraph*{}Once we have established a suitable performance of the model, we allow the model to run predictions on the un-labeled chaotic synthesizer wave forms. We output the prediction results to a file, and compare the neural network predictions against human predictions. If further corrections are needed, we revert and re-deign the features, architecture, or hyper-parameters as needed.
